---
title: "Project 3"
output:
  pdf_document: default
  html_document: default
  word_document: default
date: "October 4th 2020"
theme: cerulean
---


## Name: Project 3 

### This project deals with ways of handling string (combining and spliting them), handling factors(arranging them alphabetically and by levels), Times and dates and working with forcats.

### Splittig my name with dots.
```{r}
strsplit("Patrick.Senyo.Mensah", split = "[.]")

### This can also be separated by any letter in the name. Eg. Splitting my name with the 'y'
strsplit("Patrick.Senyo.Mensah", split = "y")
 
### with 'e'
strsplit("Patrick.Senyo.Mensah", split = "e")
```

### Combining my name with with the #paste function.
```{r}
paste("Patrick",  "Senyo", "Mensah")

### By separating them with sep.
paste("Patrick",  "Senyo", "Mensah", sep = "_")
```

### The string package is contained in tidyverse

```{r}
library(tidyverse)
### Combining names using str_c in tidyverse. str_c is vectorized.
str_c("Philant", c("Sammy", "Clement", "Wisdom", "Ephraim"), "Osman", sep = "_")
str_c("Philant", c("Sammy", "Clement", "Wisdom", "Ephraim"), c("Nash", "Osaman","Vicent"), "Osman", sep="_")

### Collapsing the above to a single string.
str_c("Philant", c("Sammy", "Clement", "Wisdom", "Ephraim"), "Osman", collapse = "")
```

### Finding the length of a string using the str_lenth function.
```{r}
### Finding the first seven characters from Philantropist.
str_sub("Philantropist", 1,7)
### Counting from backwards by finding the last seven characters from Philantropist.
str_sub("Philantropist", -7,-1)
```

### Splitting functions with str_split. This does the same thing as strsplit, execept that str_split gives you the lists. 
```{r}
str_split("www.patrickmensah.com", pattern = "[.]")
str_split(c("My full name is Patrick Senyo Mensah", "My nickname is Philant"), pattern = " ")
### Adding options.
str_split(c("My full name is Patrick Senyo Mensah", "My nickname is Philant"), pattern = " ", simplify = TRUE)
```

### Splitting strings with the boundry function. Here, each charater is separated by each 'word'
```{r}
str_split("My full name is Patrick Senyo Mensah. My nickname is Philant", boundary("word"))[[1]]
```

### Sorting strings. 
```{r}
### Arranging alphabetically.
str_sort(c("Phialnt", "Osman", "Osaman", "Clement", "Wisdom"))
### Arranging by the order of the number.
str_order(c("Phialnt", "Osman", "Osaman", "Clement", "Wisdom"))
```


## Factors
```{r}
### Days of the week in factors.
days<- c("Monday", "Tuesday","Sunday", "Wednesday", "Thursday", "Friday", "Saturday","Thursday")
### Sorting the days alphabetically.
sort(days)
### Sorting by using factors.
week.levels<- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
week<- factor(days, levels = week.levels)
sort(week)
```

```{r}
### Months of the year.
x1<- c("Mar", "Jan", "Oct", "Apr")
### Sorting the months
sort(x1)
### Arranging according to the months in the year using the levels below.
month.levels<-c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
month<-factor(x1, levels = month.levels)
sort(month)
```

## Time and dates
```{r}
### Date and times only uses lubridate.
### time only uses hms
###install.packages("lubridate")
###install.packages("hms")
```

```{r}
### loading the packages
### y means year, m means month and d means day.
library(lubridate)
library(hms)
mdy("October 4th, 2020")
dmy("4-October-2020")
ymd(20201004)
```

```{r}
### Adding time to the dates
### with hour, minute and second
ymd_hms("2020-10-04 23:14:50")
### with hour and minute
ymd_hm("2020-10-04 23:14")

### specifying the dates and time with timezones
ymd(20201004, tz= "GMT")
ymd(20201004, tz= "Asia/Singapore")
ymd(20201004, tz= "America/New_York")
```

### Getting dates and time from separate variables using make_date() and make_datetime().
```{r}
library(tidyverse)
library(lubridate)
```

```{r}
### Working with New York City flights dataset already in r.
library(nycflights13)
flights %>% select(year, month, day, hour, minute)
flights1<- flights %>% select(year, month, day, hour, minute)
### Creating a new variable 'departure' from  the above five variables.
flights1 %>% mutate(departure= make_datetime(year, month, day, hour, minute))
```

```{r}
### Creating a frequency plot for the number of departing flights 1.
flights1 %>% mutate(departure= make_datetime(year, month, day, hour, minute)) %>% ggplot(aes(departure)) + geom_freqpoly(binwidth=86400) #86400 seconds= 1 day
```

```{r}
### Keeping only flights with departure times earlier than 20130102.
flights1 %>% mutate(departure= make_datetime(year, month, day, hour, minute)) %>% filter(departure< ymd(20130102)) %>% ggplot(aes(departure)) + geom_freqpoly(binwidth=600) # 600 seconds = 10 minutes.
```

```{r}
### Keeping only flights with departure times after 20130102.
flights1 %>% mutate(departure= make_datetime(year, month, day, hour, minute)) %>% filter(departure> ymd(20130102)) %>% ggplot(aes(departure)) + geom_freqpoly(binwidth=600) # 600 seconds = 10 minutes.
```

```{r}
### Extracting date and time. 
as_date(390000)
as_datetime(390000)
datetime<-ymd_hms("3037-10-04 12:24:54")
datetime
month(datetime)
mday(datetime)
yday(datetime)
wday(datetime)
wday(datetime)
### Showing by word
wday(datetime, label = TRUE)
### with no abbreviation. 
wday(datetime, label = TRUE, abbr = FALSE)

### Converting weeks to seconds.
dweeks(12)
### months to seconds
dmonths(12)
### years to seconds
dyears(14)

### Addition and multiplication of duration
### They all covert to seconds
dyears(4) + dweeks(5)
dyears(4) + dweeks(5) +ddays(6)
3*dyears(4)

### tomorrow
tomorrow<- today()+ddays(1)
tomorrow
### last year
last.year<- today()-dyears(1)
last.year
### Next five years
next.5.years<-today()+dyears(5)
next.5.years

## Dealing with periods.
one.pm<- ymd_hms("2016-03-12 13:00:00", tz="America/New_York")
one.pm
one.pm + ddays(1)
one.pm + days(1)

### Intervals
dyears(1)/ddays(1)
```

```{r}
## Working with forcats which is a datase in r. 
library(forcats)
gss_cat
```

```{r}
### Counting by race.
gss_cat%>% count(race)
### Creating a bargraph for the count by race.
ggplot(gss_cat,aes(race)) + geom_bar()
```

```{r}
### Grouping by religion to compute the mean age and number of hours spent on watching tv.
religion.summary<- gss_cat %>% group_by(relig) %>% summarise(age=mean(age, na.rm = TRUE), tvhours=mean(tvhours, na.rm = TRUE), n=n())
religion.summary

### Making a plot with the above data.
ggplot(religion.summary, aes(tvhours, relig)) + geom_point()
```

```{r}
### Reordering the religion by the number of tv hours.
religion.summary %>% mutate(relig=fct_reorder(relig, tvhours)) %>% ggplot(aes(tvhours, relig)) + geom_point()
```

```{r}
### Reordering the levels as a factor.
levels(gss_cat$rincome)
levels(gss_cat$marital)
levels(gss_cat$race)
```

```{r}
### Reordering for a particular feedback to appear first.
levels(fct_relevel(gss_cat$rincome, "Not applicable"))
```

```{r}
### Other reordering.
marital.by.age<- gss_cat %>% filter(!is.na(age)) %>% count(age, marital) %>% group_by(age) %>% mutate(prop= n / sum(n))
marital.by.age
### Creating a plot with the above line.
ggplot(marital.by.age,aes(age, prop, colour=marital)) + geom_line(na.rm = TRUE)
```

```{r}
### Reordering levels by frequency with fct_infreq().
gss_cat %>% mutate(marital=marital %>% fct_infreq() %>% fct_rev()) %>% ggplot(aes(marital)) + geom_bar()
```

```{r}
### Collasping levels with fct_collapse().
gss_cat %>% mutate(partyid=fct_collapse(partyid, Rep = c("Strong republican", "Not str republican"), Dem = c("Strong democrat", "Not str democrat"), Ind = c("Ind,near rep", "Independent", "Ind,near dem"), Other = c("No answer", "Don't know", "Other party"))) %>% count(partyid)
```

```{r}
### Lumping groups (combine other smaller groups) together with fct_lump.
### By retaining 5 groups in 'religion'.
gss_cat %>% mutate(relig= fct_lump(relig, n=5)) %>% count(relig, sort = TRUE)
```

